---
title: "Average Bar Graph using Cibersort Data: Bladder Cancer"
output:
  pdf_document: default
  html_notebook: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("plyr")
#install.packages("reshape2")
library(dplyr)
library(reshape2)
```

# Plot Average Bar Graph using Cibersort Data - R Markdown

* Read in output from Cibersort online (tumor samples )
* Read in output from Cibersort online (normal samples )
* Create one cibersort object for each group of interest. 
* For each group of interest
    + Plot average bar for each group

## Read in cibersort output from online web - tumor samples
```{r}
# Read in cibersort output - rows are samples, columns are immune cell types
inputBLCAtumor = read.csv("input/CIBERSORTx_Tumor_Results.txt",
                         header=TRUE, row.names=1, sep="\t")
knitr::kable(inputBLCAtumor[1:5, 1:4], caption = "Bladder Cancer Tumor")
```

## Read in cibersort output from online web - normal samples
```{r}
# Read in cibersort output - rows are samples, columns are immune cell types
inputBLCAnormal = read.csv("input/CIBERSORTx_Normal_Results.txt",
                         header=TRUE, row.names=1, sep="\t")
knitr::kable(inputBLCAnormal[1:5, 1:4], caption = "Bladder cancer Normal")
```


## Create cibersort data objects
```{r}
cibersortDataTumor <- inputBLCAtumor
cibersortDataNormal <- inputBLCAnormal
```


## Call function for each group to plot average bar graph
Output will be a set of graphs in the "outputFolderName" folder

```{r}
source("functionAvgBarPlot.R")

#calling function for BLCA Tumor
objectPlotTumor <- functionAvgBarPlot(
    cibersortData = cibersortDataTumor, #set Cibersort Data ojbect for plotting
    outputFileName = "2023_Ciber_BLCATumor_AvgBarPlot", #set output file name
    outputFolderName = "output", #Set output folder name
    title1 = "BLCA Tumor  Average Bar Plot", #Change Title 
    num1 =  22)  #Number of cell types to plot

#insert object into the Rmarkdown document
objectPlotTumor

### calling function for BLCA Tumor No-relapse
objectPlotNormal <- functionAvgBarPlot(
    cibersortData = cibersortDataNormal, #set Cibersort Data ojbect for plotting
    outputFileName = "2023_Ciber_BLCANormal_AvgBarPlot", 
    #set output file name
    outputFolderName = "output", #Set output folder name
    title1 = "BLCA Normal Average Bar Plot", 
    num1 = 22) 

#insert object into the Rmarkdown
objectPlotNormal

#Build HTML / PDF
```