---
title: "03-CibersortLineGraph"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(plyr)
library(dplyr)
library(reshape2)
#install.packages("directlabels")
```

## Cibersort based Line Graph - R Markdown

* Read in output from Cibersort online (tumor samples )
* Read in output from Cibersort online (normal samples )
* Create one cibersort object for each group of interest. 
* For each group of interest
    + Call function which calculates the means of each immune cell type
    + Cbind all means values
*  Plot a line graph of the mean values so you can see how the mean value changes across the groups.
## Read in cibersort output from online web - tumor samples
```{r}
# Read in cibersort output - rows are samples, columns are immune cell types
inputBLCAtumor = read.csv("input/CIBERSORTx_Tumor_Results.txt",
                         header=TRUE, row.names=1, sep="\t")
knitr::kable(inputBLCAtumor[1:5, 1:4], caption = "Bladder cancer Tumor")
```

## Read in cibersort output from online web - normal samples
```{r}
# Read in cibersort output - rows are samples, columns are immune cell types
inputBLCAnormal = read.csv("input/CIBERSORTx_Normal_Results.txt",
                         header=TRUE, row.names=1, sep="\t")
knitr::kable(inputBLCAnormal[1:5, 1:4], caption = "Bladder cancer Normal")
```


## Create cibersort data objects
```{r}
cibersortDataTumor <- inputBLCAtumor
cibersortDataNormal <- inputBLCAnormal
```

## For each group of interest , call function to calculate means
```{r}
# Function Definition
funcMeans <- function(cibersortData) {
    
    data3 = as.data.frame(cibersortData[,1:22]) #select only 22 cell type data
    data1 = as.data.frame(t(cibersortData[,1:22]))
    
    colnames(data1) = row.names(cibersortData)
    data1$row <- row.names(data1)
    
    #melt and reshape data 
    melted <- reshape2::melt(data1, id.vars = "row")
    melted$row <-factor(melted$row, levels=c(colnames(cibersortData)))
    
    #calculate means
    means <- plyr::ddply(melted, c("row"), summarise, mean=mean(value))
    
    data2a <- reshape2::melt(means, id.vars = "row")

    return(data2a)
}

#calling function to calclulate means - returns a data frame with 22 rows and 3 columns
tumorMeans <- funcMeans(cibersortDataTumor)
normalMeans <- funcMeans(cibersortDataNormal)

#The First column is the cell type, Third column is the mean value
colnames(tumorMeans) 
head(tumorMeans)
```

## cbind all means value (third column) from each object 
IMPORTANT NOTE - When doing cbind, order them such that the baseline group is on the left most column
```{r}
#IMPORTANT NOTE - When doing cbind, order them such that the baseline group is on the left most column
meansAll = cbind(normalMeans$value, tumorMeans$value)
row.names(meansAll) <- tumorMeans$row #Set row names - immune cell types
colnames(meansAll) <- c("Adjacent Normal", "Tumor")
```

## Call function to Plot Line Graph. 
Output will be a set of graphs in the "outputFolderName" folder

```{r}
source("functionLineGraph.R")

objectLineGraph <- functionLineGraph(meansAll, 
                                     outputFileName = "2BLCA_Ciber_LinePlot", 
                                     outputFolderName = "output") #Set output folder name

#insert object into the Rmarkdown
objectLineGraph

#Build HTML / PDF
```