---
title: "Aizhan_Step1"
author: "aijeanka"
date: "2023-11-29"
output: html_document
---
---
title: "Final Project: Step 1"
author: "Aizhan Uteubayeva (NetID:au198)"
output: html_document
---
## Step 1- Read in data into R. Check if data was imported correctly
### 1a - Read in clinical data
```{r}
library(readr)
# Reading the data into Rc.clinData
clinData <- read.csv(file = "input/20191105_BC_ClinData_233rows.csv",
                       sep=",", 
                       header = TRUE, 
                       stringsAsFactors = FALSE, 
                       row.names = 1)

# Displaying the first 6 rows and first 4 columns using knitr::kable
knitr::kable(head(clinData[, 1:4], n = 6))
```
### 1b: Read in processed normalized gene expression data in log2 scale, includes gene annotation
```{r}
geneExp <- read.csv(file = "input/20191105_BC_GeneExpData_withAnno_233.tsv",
                       sep="\t", 
                       header = T, 
                       stringsAsFactors = F)

knitr::kable(geneExp[1:5, 1:4], caption = "Processed gene expression data in log2 scale with gene annotation")
```

### The Goal is to compare which two groups ?
Team 1 goal is to compare Pre-Cancer (normal looking bladder mucosae surrounding cancer) vs Normal patients 
-   Question: What is the outcome variable column from clinical data file ? It is the column in the clinical file that contains the baseline and comparison values
-   Answer: "PrimaryBladderCancerType" column has values "Normal bladder mucosae" (baseline group) and "Bladder muscosae surrounding cancer" (comparison group)
```{r}
unique_types <- unique(clinData$PrimaryBladderCancerType)
print(unique_types)
# in Team 1 we are focusing on "Bladder muscosae surrounding cancer" and "Normal bladder mucosae"
```

## Step 2 - Clean/Filter data

### Filter clinical data 
#### 1(a) Create the dataframe with baseline patients, leaving only ID and "Normal bladder mucosae"
```{r}
# Looking at the clinData dataframe features
dim(clinData)
colnames(clinData)
summary(clinData)
```
#### Looking at the baseline patients
There are 10 patients with "Normal bladder mucosae"
```{r}
library(dplyr)

baseline <- clinData %>%
  filter(PrimaryBladderCancerType == "Normal bladder mucosae") %>%
  select(GSMid, PrimaryBladderCancerType)

head(baseline)
``` 
#### 1(b) Create the dataframe with comparison patients, leaving only ID and "Normal bladder mucosae"
There are 58 patients with "Bladder mucosae surrounding cancer"
```{r}
precancer <- clinData %>%
  filter(PrimaryBladderCancerType == "Bladder mucosae surrounding cancer") %>%
  select(GSMid, PrimaryBladderCancerType)

head(precancer)
``` 
### Cleaning gene expression data
Check rows and columns:

-   Rows are features (genes in this case)
-   Columns are patients

```{r}
knitr::kable(geneExp[1:5, 1:4]) 
```

#### Now creating gene expression data for tumor samples
The GSM ids are the identifiers in the gene exp data that match the identifiers for the Clinical data file:

-   Sample identifier for ClinData: Column called “GSMid”
-   Sample identifier for geneExp: Column names are the IDs
```{r}
matchingSamples <- which(colnames(geneExp) %in% clinData$GSMid) # 233 IDs matching
subsetGeneExp <- geneExp[, matchingSamples] ## 431348 rows features, 233 patient columns
``` 

## Step 3 - Identify the groups to be compared

### Identify the groups to be compared (Baseline and Comparison Grps)
```{r}
# Labels (row numbers) that can identify the baseline group patients
baselineGrpLabels <-  which(clinData$PrimaryBladderCancerType == "Normal bladder mucosae") 
head(baselineGrpLabels)
# Use the labels (row numbers) to subset baseline patients in clinical data file = 10 patients
clinBase <- clinData[baselineGrpLabels, ] # we have got our 10 patients for baseline

# Labels (row numbers) that can identify the comp group patients = 58 patients
compGrpLabels <- which(clinData$PrimaryBladderCancerType == "Bladder mucosae surrounding cancer") 
head(compGrpLabels)
# Use the labels (row numbers) to subset comp patients in clinical data file 
clinComp <- clinData[compGrpLabels, ] # got 58 patients for pre-cancer

# Use the clinBase and clinComp objects to subset gene expression data
geneExpTumorBase <- subsetGeneExp[, clinBase$GSMid] # 43148 feature (rows), 10 samples columns
geneExpTumorComp <- subsetGeneExp[, clinComp$GSMid] # 43148 feature (rows), 58 samples columns
```

## Step 4- Sanity check

(Yes)  See if filtering of clinical data in R matches filtering of clinical data in excel
(Yes)  See if sample ids in clinical data match sample ids in gene exp data (if they don't match it means your step 1 and/or 2 is wrong)
(Yes)  Verify you see correct number of samples in baseline and comp groups
(Yes)  Export the column names from gene expression data to see if it contains only probe/gene names and no other garbage

```{r}
#See if sample ids in clinical data match sample ids in gene exp data
clinBase$GSMid == colnames(geneExpTumorBase)
clinComp$GSMid == colnames(geneExpTumorComp)

### Checking if the samples IDs baseline and comp groups are the same ---- you can put these in an excel to see if IDs match
write.table(clinBase$GSMid , file = "output/ClinBaseIDs.tsv", sep="\t", quote = F  )
write.table(clinComp$GSMid , file = "output/ClinCompIDs.tsv", sep="\t", quote = F  )
write.table(colnames(geneExpTumorBase) , 
            file = "output/GeneExpBaseIDs.tsv", 
            sep="\t", quote = F  )
write.table(colnames(geneExpTumorComp) , 
            file = "output/GeneExpCompIDs.tsv", 
            sep="\t", quote = F  )

#Export the features from gene expression data  
#Open this file and check that it contains only probe/gene names and no other garbage
write.table(rownames(geneExp),file = "output/genExpFeatureIDsCheck.tsv", sep="\t", quote = F  )
```
## Step 5 - Prep data for T-test
```{r}
### Checking to make sure data is a numeric data frame
knitr::kable(head(geneExpTumorBase[1:5,1:4]))
knitr::kable(head(geneExpTumorComp[1:5,1:4]))
```
#### Function for T-test
```{r ttest}
source("fnTTest.R")

#### Call T-test function
results1 = fnTTest(baseGroup = geneExpTumorBase, 
                   compGroup = geneExpTumorComp, 
                   testName = "Aizhan_Bladder_TTest", 
                   baseGroupName = "Normal Mucosae",
                   compGroupName =  "Pre-Cancer Mucosae",
                   folderName = "output")
```
```{r}
sum(is.na.data.frame(geneExpTumorComp))
sum(is.na.data.frame(geneExpTumorBase))
```
## Brushing up
### Creating two csv finles with ordered p-value>0.001 and top 20 genes
```{r}
ttest <- read.csv(file = "output/Aizhan_Bladder_TTest.csv")

#check to make sure p-value column is imported as numeric 
#sort by p-value (just in case the results are not sorted by p-value)
ttestResultsSorted <- dplyr::arrange(ttest, Pvalue)

#find rows with p-value < 0.001
whichSig <- which(ttestResultsSorted$Pvalue <= 0.001) 

#Short list sig results
ttestResultsSig <- unique(ttestResultsSorted[whichSig, ]) #821 rows

### Export short listed results
write.table(x = ttestResultsSig, 
            file = "output/Aizhan_Bladder_Ttest_Shortlisted.tsv", 
            quote = F, sep = "\t")
```

