---
title: "Final Project: Step 3"
author: "Aizhan Uteubayeva (NetID:au198)"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(plyr)
library(dplyr)
library(reshape2)
library(tibble)

# install.packages("directlabels")
```

## Reading in the input data

### Normal Tissue data

```{r}
# Read in Cibersort Normal output - rows are samples, columns are immune cell types
NormalBladder = read.csv("CIBERSORTx_Normal_Results.csv",
                         header=TRUE, row.names=1, sep="\t")

# Displaying the data
knitr::kable(NormalBladder[1:5, 1:4], caption = "Normal Bladder Mucosa")
```

### Tumor Tissue data for Surrounding Mucosa

Fractions for 22 types and three columns with additional statistics

```{r}
# Read in cibersort Tumor output - rows are samples, columns are immune cell types
TumorBladder = read.csv("CIBERSORTx_Tumour_Results.csv",
                         header=TRUE, row.names=1, sep="\t")
knitr::kable(TumorBladder[1:5, 1:4], caption = "Colon cancer Tumor")
```

### Cleaning the values

```{r}
NormalCleaned <- NormalBladder[, -which(names(NormalBladder) %in% c("Correlation", "RMSE","P.value"))]
TumorCleaned <- TumorBladder[, -which(names(TumorBladder) %in% c("Correlation", "RMSE","P.value"))]
```

## Top five in normal cells

```{r}
normal_means <- colMeans(NormalCleaned)
normal_means_df <- data.frame(normal_means)
normal_means <- tibble::rownames_to_column(normal_means_df, "VALUE")


top_normal <- normal_means[order(normal_means_df$normal_means, decreasing = TRUE),]

top_5_normal <- top_normal[1:5, ]
knitr::kable(top_5_normal, caption = "Top five cells in Normal Bladder Surrounding Mucosa")
```

## Top five in Surrounding Mucosa

```{r}
tumor_means <- colMeans(TumorCleaned)
tumor_means_df <- data.frame(tumor_means)

tumor_means <- tibble::rownames_to_column(tumor_means_df, "VALUE")
top_tumor <- tumor_means[order(tumor_means_df$tumor_means, decreasing = TRUE),]

top_5_tumor <- top_tumor[1:5, ]
knitr::kable(top_5_tumor, caption = "Top five cells in Pre-tumor Bladder Surrounding Mucosa")
```

## Cibersort

#### Creating cibersort data objects
```{r}
cibersortDataTumor <- TumorBladder
cibersortDataNormal <- NormalBladder
```
### Plotting the Average Bar Graph
```{r}
source("functionAvgBarPlot.R")

#calling function for CRC Tumor
objectPlotTumor <- functionAvgBarPlot(
    cibersortData = cibersortDataTumor, #set Cibersort Data ojbect for plotting
    outputFileName = "2023_Ciber_CRCTumor_AvgBarPlot", #set output file name
    outputFolderName = "output", #Set output folder name
    title1 = "CRC Tumor  Average Bar Plot", #Change Title 
    num1 =  22)  #Number of cell types to plot

#insert object into the Rmarkdown document
objectPlotTumor

### calling function for CRC Tumor No-relapse
objectPlotNormal <- functionAvgBarPlot(
    cibersortData = cibersortDataNormal, #set Cibersort Data ojbect for plotting
    outputFileName = "2023_Ciber_CRCNormal_AvgBarPlot", 
    #set output file name
    outputFolderName = "output", #Set output folder name
    title1 = "CRC Normal Average Bar Plot", 
    num1 = 22) 

#insert object into the Rmarkdown
objectPlotNormal

#Build HTML / PDF
```

### Plotting Stacked Bar Graph
```{r}
source("functionStackedBarPlot.R")

#calling function for CRC Tumor 
objectPlotT <- functionStackedBarPlot(
    cibersortData = cibersortDataTumor, #set Cibersort Data ojbect for plotting
    outputFileName = "2022_Ciber_CRCTumor_StackedBarPlot", #set output file name
    outputFolderName = "output", #Set output folder name
    title1 = "CRC Tumor Stacked Bar Plot") #Change Title

#insert object into the Rmarkdown
objectPlotT

### calling function for CRC normal
objectPlotN <- functionStackedBarPlot(
    cibersortData = cibersortDataNormal, #set Cibersort Data ojbect for plotting
    outputFileName = "2022_Ciber_CRCNormal_StackedBarPlot", 
    #set output file name
    outputFolderName = "output", #Set output folder name
    title1 = "CRC Normal Stacked Bar Plot") #Change Title

#insert object into the Rmarkdown
objectPlotN

#Build HTML / PDF
```

#### Calculating means for each group
```{r}
# Function Definition
funcMeans <- function(cibersortData) {
    
    data3 = as.data.frame(cibersortData[,1:22]) #select only 22 cell type data
    data1 = as.data.frame(t(cibersortData[,1:22]))
    
    colnames(data1) = row.names(cibersortData)
    data1$row <- row.names(data1)
    
    #melt and reshape data 
    melted <- reshape2::melt(data1, id.vars = "row")
    melted$row <-factor(melted$row, levels=c(colnames(cibersortData)))
    
    #calculate means
    means <- plyr::ddply(melted, c("row"), summarise, mean=mean(value))
    
    data2a <- reshape2::melt(means, id.vars = "row")

    return(data2a)
}

#calling function to calclulate means - returns a data frame with 22 rows and 3 columns
tumorMeans <- funcMeans(cibersortDataTumor)
normalMeans <- funcMeans(cibersortDataNormal)

#The First column is the cell type, Third column is the mean value
colnames(tumorMeans) 
head(tumorMeans)
```

#### cbind all means value (third column) from each object

```{r}
#IMPORTANT NOTE - When doing cbind, order them such that the baseline group is on the left most column
meansAll = cbind(normalMeans$value, tumorMeans$value)
row.names(meansAll) <- tumorMeans$row #Set row names - immune cell types
colnames(meansAll) <- c("Adjacent Normal", "Tumor")
```

### Plotting Line Graph

```{r}
source("functionLineGraph.R")

objectLineGraph <- functionLineGraph(meansAll, 
                                     outputFileName = "2023_Ciber_LinePlot", 
                                     outputFolderName = "output") #Set output folder name

#insert object into the Rmarkdown
objectLineGraph

#Build HTML / PDF
```
